Plots

### 2025-10-12 src/plots.Rmd
Workng on gpt5 plots

# Plot performance of models using lines and points
plot_perf <- function(
        df_long,
        f = calc_pccc,
        group = "cod_name",
        color = "model_name",
        text_scale = 1,
        title = NULL,
        title_size = 10,
        x_title = "Partial Chance Corrected Concordance (PCCC)",
        y_title = NULL,
        text_mid_size = 2.25 * text_scale,
        text_mid_vjust = 2.5,
        text_start_size = 2.75 * text_scale,
        text_end_size = 3.5 * text_scale,
        text_hjust = 0.2,
        add_min_points = F,
        add_max_points = F,
        mid_points_size = 4,
        min_points_size = 4,
        max_points_size = 4,
        x_limits = c(0, 1.15),
        x_breaks = seq(0, 1, by = 0.1),
        x_expand = expansion(add = 0.2),
        x_color_labels = c(
            "G3" = "GPT-3.5 (G3)",
            "G4" = "GPT-4 (G4)",
            "ISV" = "InSilicoVA (ISV)",
            "I5" = "InterVA-5 (I5)"
        ),
        x_shapes = 21:25,
        x_title_size = 9 * text_scale,
        y_title_size = 9 * text_scale,
        x_text_size = 8 * text_scale,
        y_text_size = 9 * text_scale,
        sort_group = T,
        legend_position = "bottom",
        legend_title = element_blank(),
        legend_margin = margin(c(-8, -2, -2, -50)),
        legend_add = c("R = Range"),
        ...
    ) {

@@

# Plot perf of models using trend lines
plot_perf_trend <- function(
        df_long,
        f = calc_pccc,
        group = "age_range",
        text_scale = 1,
        title = NULL,
        x_title = NULL,
        y_title = NULL,
        title_size = 10,
        x_title_size = 9 * text_scale,
        y_title_size = 9 * text_scale,
        x_text_size = 8 * text_scale,
        y_text_size = 9 * text_scale,
        legend_text_size = 9 * text_scale,
        legend_position = "bottom",
        legend_direction = "horizontal",
        legend_spacing = unit(0, "pt"),
        x_color_labels = c(
            "G3" = "GPT-3.5",
            "G4" = "GPT-4",
            "ISV" = "InSilicoVA",
            "I5" = "InterVA-5"
        ),
        x_linetypes = c(
            "solid",
            "solid",
            "dashed",
            "dotted"
        ),
        plot_mean = F
    ) {

@@

# Read preds
df_raw <- read_csv("../data/healsl_rd1to2_cod_v1.csv")

@@

## Combined Models

Create model combinations where if any of the models have the physician code, then set the combined model's output for the case to be the physician code.

# Calc combine preds
df <- df %>%
    mutate(
        "gpt3_insilicova_cghr10" = if_else(
            gpt3_cghr10 == physician_cghr10 | 
                insilicova_cghr10 == physician_cghr10,
            physician_cghr10,
            gpt3_cghr10
        ),
        "gpt4_insilicova_cghr10" = if_else(
            gpt4_cghr10 == physician_cghr10 | 
                insilicova_cghr10 == physician_cghr10,
            physician_cghr10,
            gpt4_cghr10
        ),
        "gpt3_interva5_cghr10" = if_else(
            gpt3_cghr10 == physician_cghr10 | 
                interva5_cghr10 == physician_cghr10,
            physician_cghr10,
            gpt3_cghr10
        ),
        "gpt4_interva5_cghr10" = if_else(
            gpt4_cghr10 == physician_cghr10 | 
                interva5_cghr10 == physician_cghr10,
            physician_cghr10,
            gpt4_cghr10
        ),
        "gpt3_insilicova_interva5_cghr10" = if_else(
            gpt3_cghr10 == physician_cghr10 | 
                insilicova_cghr10 == physician_cghr10 |
                interva5_cghr10 == physician_cghr10,
            physician_cghr10,
            gpt3_cghr10
        ),
        "gpt4_insilicova_interva5_cghr10" = if_else(
            gpt4_cghr10 == physician_cghr10 | 
                insilicova_cghr10 == physician_cghr10 |
                interva5_cghr10 == physician_cghr10,
            physician_cghr10,
            gpt4_cghr10
        )
    )

# Display combine preds
print(head(df %>% select(
    gpt3_insilicova_cghr10,
    gpt4_insilicova_cghr10,
    gpt3_interva5_cghr10,
    gpt4_interva5_cghr10,
    gpt3_insilicova_interva5_cghr10,
    gpt4_insilicova_interva5_cghr10
)))

@@

# Rename COD names
df <- df %>%
    mutate(across(
        c(
            physician_cghr10,
            gpt3_cghr10,
            gpt4_cghr10,
            insilicova_cghr10,
            interva5_cghr10
        ),
        ~ case_when(
            .x == "Maternal conditions" ~ "Maternal",
            .x == "Acute respiratory infections" ~ "Pneumonia",
            .x == "Neonatal infections" ~ "Infections",
            .x == "Road and transport injuries" ~ "Road/transport injuries",
            .x == "Liver and alcohol related diseases" ~ "Liver/alcohol",
            .x == "Birth asphyxia and trauma" ~ "Birth asphyxia/trauma",
            .default = .x
        ) %>%
            str_replace_all(" diseases| disease", "")
    ))

@@

# Shorten COD names
df <- df %>%
    mutate(across(
        c(
            physician_cghr10,
            gpt3_cghr10,
            gpt4_cghr10,
            insilicova_cghr10,
            interva5_cghr10
        ),
        ~ str_wrap(.x, width = 18)
    ))

@@

# Model full name
gpt3_name <- "G3"
gpt4_name <- "G4"
isva_name <- "ISV"
iva5_name <- "I5"

# Pivot data into long format for calc
df_long <- df %>%
    filter( # remove low count cod
        physician_cghr10_n >= 10
    ) %>%
    pivot_longer( # long format
        ends_with("_cghr10") & !starts_with("physician_"),
        names_to = "model",
        values_to = "model_cghr10"
    ) %>%
    mutate(
        model = str_replace( # remove cghr10
            model,
            "_cghr10",
            ""
        ),
        model_name = str_replace_all(model, "_", " or ") %>% # use full model names
            str_replace(., "gpt3", gpt3_name) %>%
            str_replace(., "gpt4", gpt4_name) %>%
            str_replace(., "insilicova", isva_name) %>%
            str_replace(., "interva5", iva5_name) %>%
            factor(., levels = c(
                gpt4_name,
                gpt3_name,
                isva_name,
                iva5_name
            )),
        age_name = case_when( # use full age name
            age == "adult" ~ "Adult",
            age == "child" ~ "Child",
            age == "neo" ~ "Neonate",
            .default = NA
        ),
        cod = physician_cghr10,
        cod_name = paste0(
            cod,
            "\n(n=", physician_cghr10_n, ", ", round(physician_cghr10_perc), "%)"
        ),
        ncod = case_when( # num of cods for each age group
            age == "adult" ~ ncod_adult,
            age == "child" ~ ncod_child,
            age == "neo" ~ ncod_neo,
            .default = NA
        )
    ) %>%
    filter( # remove combine models
        !str_detect(model_name, " or ")
    )

@@

# Prep data
df_plot <- df_long %>%
    group_by(model_name) %>%
    summarize(
        pccc = calc_pccc(cod, model_cghr10, N = ncod_all),
        csmf = calc_csmf_acc(cod, model_cghr10)
    ) %>%
    arrange(pccc) %>%
    mutate(
        model_name = model_name %>%
            case_when(
                . == "G3" ~ "GPT-3.5",
                . == "G4" ~ "GPT-4",
                . == "ISV" ~ "InSilicoVA",
                . == "I5" ~ "InterVA-5",
                .default = .
            ) %>%
            factor(., levels = .)
    ) %>%
    pivot_longer(
        c(pccc, csmf),
        names_to = "metric",
        values_to = "metric_value"
    ) %>%
    mutate(
        metric = str_to_upper(metric) %>%
            str_replace(., "CSMF", "Cause Specific Mortality Fraction (CSMF) Accuracy") %>%
            str_replace(., "PCCC", "Partial Chance Corrected Concordance (PCCC)") %>%
            factor(., levels = rev(c("Partial Chance Corrected Concordance (PCCC)", "Cause Specific Mortality Fraction (CSMF) Accuracy"))),
        metric_label = paste0(metric, "=", round(metric_value, 2)) %>%
            str_replace(., " Accuracy", ""),
        metric_pattern = case_when(
            metric == "Partial Chance Corrected Concordance (PCCC)" ~ "stripe",
            metric == "Cause Specific Mortality Fraction (CSMF) Accuracy" ~ "none"
        )
    )

>>>>

# Plot performance of models using lines and points
plot_perf <- function(
        df_long,
        f = calc_pccc,
        group = "cod_name",
        color = "model_name",
        text_scale = 1,
        title = NULL,
        title_size = 10,
        x_title = "Partial Chance Corrected Concordance (PCCC)",
        y_title = NULL,
        text_mid_size = 2.25 * text_scale,
        text_mid_vjust = 2.5,
        text_start_size = 2.75 * text_scale,
        text_end_size = 3.5 * text_scale,
        text_hjust = 0.2,
        add_min_points = F,
        add_max_points = F,
        mid_points_size = 4,
        min_points_size = 4,
        max_points_size = 4,
        x_limits = c(0, 1.15),
        x_breaks = seq(0, 1, by = 0.1),
        x_expand = expansion(add = 0.2),
        x_color_labels = c(
            "G3" = "GPT-3.5 (G3)",
            "G4" = "GPT-4 (G4)",
            "G5" = "GPT-5 (G5)",
            "ISV" = "InSilicoVA (ISV)",
            "I5" = "InterVA-5 (I5)"
        ),
        x_shapes = 21:26,
        x_title_size = 9 * text_scale,
        y_title_size = 9 * text_scale,
        x_text_size = 8 * text_scale,
        y_text_size = 9 * text_scale,
        sort_group = T,
        legend_position = "bottom",
        legend_title = element_blank(),
        legend_margin = margin(c(-8, -2, -2, -50)),
        legend_add = c("R = Range"),
        ...
    ) {

@@

# Plot perf of models using trend lines
plot_perf_trend <- function(
        df_long,
        f = calc_pccc,
        group = "age_range",
        text_scale = 1,
        title = NULL,
        x_title = NULL,
        y_title = NULL,
        title_size = 10,
        x_title_size = 9 * text_scale,
        y_title_size = 9 * text_scale,
        x_text_size = 8 * text_scale,
        y_text_size = 9 * text_scale,
        legend_text_size = 9 * text_scale,
        legend_position = "bottom",
        legend_direction = "horizontal",
        legend_spacing = unit(0, "pt"),
        x_color_labels = c(
            "G3" = "GPT-3.5",
            "G4" = "GPT-4",
            "G5" = "GPT-5",
            "ISV" = "InSilicoVA",
            "I5" = "InterVA-5"
        ),
        x_linetypes = c(
            "solid",
            "solid",
            "dashed",
            "dotted",
            "dotted"
        ),
        plot_mean = F
    ) {

@@

# Read preds
df_raw <- read_csv("../data/healsl_rd1to2_cod_v2.csv")

@@

@@

# Rename COD names
df <- df %>%
    mutate(across(
        c(
            physician_cghr10,
            gpt3_cghr10,
            gpt4_cghr10,
            gpt5_cghr10,
            insilicova_cghr10,
            interva5_cghr10
        ),
        ~ case_when(
            .x == "Maternal conditions" ~ "Maternal",
            .x == "Acute respiratory infections" ~ "Pneumonia",
            .x == "Neonatal infections" ~ "Infections",
            .x == "Road and transport injuries" ~ "Road/transport injuries",
            .x == "Liver and alcohol related diseases" ~ "Liver/alcohol",
            .x == "Birth asphyxia and trauma" ~ "Birth asphyxia/trauma",
            .default = .x
        ) %>%
            str_replace_all(" diseases| disease", "")
    ))

@@

# Shorten COD names
df <- df %>%
    mutate(across(
        c(
            physician_cghr10,
            gpt3_cghr10,
            gpt4_cghr10,
            gpt5_cghr10,
            insilicova_cghr10,
            interva5_cghr10
        ),
        ~ str_wrap(.x, width = 18)
    ))

@@

# Model full name
gpt3_name <- "G3"
gpt4_name <- "G4"
gpt5_name <- "G5"
isva_name <- "ISV"
iva5_name <- "I5"

# Pivot data into long format for calc
df_long <- df %>%
    filter( # remove low count cod
        physician_cghr10_n >= 10
    ) %>%
    pivot_longer( # long format
        ends_with("_cghr10") & !starts_with("physician_"),
        names_to = "model",
        values_to = "model_cghr10"
    ) %>%
    mutate(
        model = str_replace( # remove cghr10
            model,
            "_cghr10",
            ""
        ),
        model_name = str_replace_all(model, "_", " or ") %>% # use full model names
            str_replace(., "gpt3", gpt3_name) %>%
            str_replace(., "gpt4", gpt4_name) %>%
            str_replace(., "gpt5", gpt5_name) %>%
            str_replace(., "insilicova", isva_name) %>%
            str_replace(., "interva5", iva5_name) %>%
            factor(., levels = c(
                gpt5_name,
                gpt4_name,
                gpt3_name,
                isva_name,
                iva5_name
            )),
        age_name = case_when( # use full age name
            age == "adult" ~ "Adult",
            age == "child" ~ "Child",
            age == "neo" ~ "Neonate",
            .default = NA
        ),
        cod = physician_cghr10,
        cod_name = paste0(
            cod,
            "\n(n=", physician_cghr10_n, ", ", round(physician_cghr10_perc), "%)"
        ),
        ncod = case_when( # num of cods for each age group
            age == "adult" ~ ncod_adult,
            age == "child" ~ ncod_child,
            age == "neo" ~ ncod_neo,
            .default = NA
        )
    ) %>%
    filter( # remove combine models
        !str_detect(model_name, " or ")
    )

@@

# Prep data
df_plot <- df_long %>%
    group_by(model_name) %>%
    summarize(
        pccc = calc_pccc(cod, model_cghr10, N = ncod_all),
        csmf = calc_csmf_acc(cod, model_cghr10)
    ) %>%
    arrange(pccc) %>%
    mutate(
        model_name = model_name %>%
            case_when(
                . == "G3" ~ "GPT-3.5",
                . == "G4" ~ "GPT-4",
                . == "G5" ~ "GPT-5",
                . == "ISV" ~ "InSilicoVA",
                . == "I5" ~ "InterVA-5",
                .default = .
            ) %>%
            factor(., levels = .)
    ) %>%
    pivot_longer(
        c(pccc, csmf),
        names_to = "metric",
        values_to = "metric_value"
    ) %>%
    mutate(
        metric = str_to_upper(metric) %>%
            str_replace(., "CSMF", "Cause Specific Mortality Fraction (CSMF) Accuracy") %>%
            str_replace(., "PCCC", "Partial Chance Corrected Concordance (PCCC)") %>%
            factor(., levels = rev(c("Partial Chance Corrected Concordance (PCCC)", "Cause Specific Mortality Fraction (CSMF) Accuracy"))),
        metric_label = paste0(metric, "=", round(metric_value, 2)) %>%
            str_replace(., " Accuracy", ""),
        metric_pattern = case_when(
            metric == "Partial Chance Corrected Concordance (PCCC)" ~ "stripe",
            metric == "Cause Specific Mortality Fraction (CSMF) Accuracy" ~ "none"
        )
    )
